Evaluating machine translation for assimilation via a gap-filling task

### 2015-04-15 2015-eamt-assim/2015-eamt-assim.tex
Minor edits
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59840 72bbbca6-d526-0410-a7d9-f06f51895060

\maketitle
\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
\begin{abstract}
  This paper provides additional observations on the viability of a
  strategy independently proposed in 2012 and 2013 for evaluation of
 MT for assimilation purposes. The strategy involves
  human evaluators, who are asked to restore keywords (to fill gaps) in
  reference translations. The evaluation method is applied to two language pairs, Basque--Spanish and Tatar--Russian. To reduce the amount of time required to prepare tasks
  and analyse results, an open-source task management system is
  introduced. The evaluation results show that the gap-filling task may be suitable for measuring MT
  quality for assimilation purposes.
\end{abstract}

@@

Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations\footnote{\url{http://googleblog.blogspot.co.uk/2012/04/breaking-down-language-barriersix-years.html}}), few methodologies are established for assimilation evaluation of MT. The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}. These approaches are often costly and prone to subjectivity: see the discussion in \cite{oregan13}. As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method. Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}. While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation. A designated number of keywords is removed from the human-translated sentences. The evaluators are then asked to fill the gaps with suitable words. The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions. Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.

@@

The paper is organised as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated. Section~\ref{sec:setup} introduces the experimental material, the evaluators, the distribution of tasks and the evaluation procedure. In section~\ref{sec:results} we describe and discuss the experiment results. Finally, section~\ref{sec:conclusion} draws some conclusions. This paper is concerned primarily with assimilation evaluation; for a deeper discussion on evaluation see e.g. \citep{koehn2009}.

@@

For each experiment we called for native speakers of target language of the language pair (i.e.
Spanish and Russian) who had no command of source language of the pair (Basque
and Tatar, respectively). The knowledge was self-reported, and the participants were not asked about any other languages they may know. \comment{EA 12: should I elaborate on this? Write that Basque is not mutually understandable with anything? I am pretty sure we had no one from Tatarstan, but there may have been someone who knows some Bashkir, which might have helped them. Also, with Tatar, I think, Russian loanwords help more than anything else. } Eleven evaluators participated in the Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
in full, see discussion). The majority of Russian participants were aged 20-25, with university degrees or in the process of obtaining them. Although we have not asked the participants about their knowledge of languages other than Tatar and Russian, it is reasonable to assume that most Russian participants knew English to some extent. \comment{EA 12: could you tell me more about the Spanish evaluators? I would write something similar}

@@

% this section talks about percentages in eu-es and tat-rus, and how everything is nice.
We will now discuss the percentage of correct answers based on task type. In general, tasks with MT hints score higher than tasks without MT hints. This aligns well with our expectations and suggests than machine translation helps to understand the provided text. In addition, tasks with source hints are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks. In view of the relatively large standard deviations, the significance of the hints' contribution was tested using a linear regression model. The data points were represented as an individual evaluator's average score (the number of correct answers divided by the total number of answers) in each of the percentage-hint combinations. Two separate models were created: one for no-hint vs MT-hint tasks, and another for no-hint vs source-hint tasks. Given the null hypothesis of the regression line slope equal to zero, the contribution of MT hint is significant on the p~<~0.001 level. The contribution of the source hint is insignificant with p~=~0.162.

@@

We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated). To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect. The previously obtained synonym lists were taken into account, i.e. if the two answers are different but both correct, they fall into one category. The measure was calculated separately for each hint and percentage combination.

The interpretation of Krippendorff's alpha varies depending on the application. One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: k < 0.00 indicates ``poor'' agreement, 0.00 to 0.20 ``slight'', 0.21 to 0.40 ``fair'', 0.41 to 0.60 ``moderate'', 0.61 to 0.80 ``substantial'', and 0.81 to 1.00 ``near perfect''.

@@

The experiments may easily be repeated for any language pair (provided a parallel corpus) and
any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
\begin{enumerate}
\item As reported in \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus. Neighboring sentences on the same topic may assist the users in gap-filling tasks;
\item If possible, increase the number of evaluators, or reduce the number of questions per participant. In the above experiments each participant filled from 110 to 187 gaps, divided into small groups. Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}\comment{EA: the participants did not leave the gaps empty, they just did not attempt the next portions of evaluation after finishing some and getting bored. or did I misunderstand the use of captchas? }
\item To account for the adaptation period, provide training tasks before the main evaluations take place.
\end{enumerate}

>>>>

\maketitle
\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
\begin{abstract}
  This paper provides additional observations on the viability of a
  strategy independently proposed in 2012 and 2013 for evaluation of
 machine translation (MT) for assimilation purposes. The strategy involves
  human evaluators, who are asked to restore keywords (to fill gaps) in
  reference translations. The evaluation method is applied to two language pairs, Basque--Spanish and Tatar--Russian. To reduce the amount of time required to prepare tasks
  and analyse results, an open-source task management system is
  introduced. The evaluation results show that the gap-filling task may be suitable for measuring MT
  quality for assimilation purposes.
\end{abstract}

@@

Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations\footnote{\url{http://googleblog.blogspot.co.uk/2012/04/breaking-down-language-barriersix-years.html}}), few methodologies are established for assimilation evaluation of MT. The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}. These approaches are often costly and prone to subjectivity: see the discussion by \cite{oregan13}. As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method. Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}. While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation. A designated number of keywords is removed from the human-translated sentences. The evaluators are then asked to fill the gaps with suitable words. The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions. Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.

@@

The paper is organised as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated. Section~\ref{sec:setup} introduces the experimental material, the evaluators, the distribution of tasks and the evaluation procedure. In section~\ref{sec:results} we describe and discuss the experiment results. Finally, section~\ref{sec:conclusion} draws some conclusions. This paper is concerned primarily with assimilation evaluation; for a deeper discussion on evaluation see e.g. \citep{koehn2010}.

@@

For each experiment we called for native speakers of target language of the language pair (i.e.
Spanish and Russian) who had no command of source language of the pair (Basque
and Tatar, respectively). The knowledge was self-reported, and the participants were not asked about any other languages they may know. \comment{EA 12: should I elaborate on this? Write that Basque is not mutually understandable with anything? I am pretty sure we had no one from Tatarstan, but there may have been someone who knows some Bashkir, which might have helped them. Also, with Tatar, I think, Russian loanwords help more than anything else. } Eleven evaluators participated in the Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
in full, see discussion). The majority of Russian participants were aged 20-25, with university degrees or in the process of obtaining them. Although we have not asked the participants about their knowledge of languages other than Tatar and Russian, it is reasonable to assume that most Russian participants knew English to some extent. The Spanish participants were university staff with background in Computer Science.

@@

% this section talks about percentages in eu-es and tat-rus, and how everything is nice.
We will now discuss the percentage of correct answers based on task type. In general, tasks with MT hints score higher than tasks without MT hints. This aligns well with our expectations and suggests than machine translation helps to understand the provided text. In addition, tasks with source hints are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks. In view of the relatively large standard deviations, the significance of the hints' contribution was tested using a linear regression model. The data points were represented as an individual evaluator's average score (the number of correct answers divided by the total number of answers) in each of the percentage-hint combinations. Two separate models were created: one for no-hint vs MT-hint tasks, and another for no-hint vs source-hint tasks. Given the null hypothesis of the regression line slope equal to zero, the contribution of MT hint is significant on the p~<~0.001 level, while the contribution of the source hint is significant only with p~<~0.162

@@

We selected this measure because of its compatibility with more than two annotators per task and missing data (not all the gaps were evaluated). To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect. The previously obtained synonym lists were taken into account, i.e. if the two answers are different but both correct, they fall into one category. The measure was calculated separately for each hint and percentage combination.

The interpretation of Krippendorff's alpha varies depending on the application. One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: k < 0 indicates ``poor'' agreement, 0 to 0.2 ``slight'', 0.21 to 0.4 ``fair'', 0.41 to 0.6 ``moderate'', 0.61 to 0.8 ``substantial'', and 0.81 to 1 ``near perfect''.

@@

The experiments may easily be repeated for any language pair (provided a parallel corpus) and
any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
\begin{enumerate}
\item As reported by \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus. Neighboring sentences on the same topic may assist the users in gap-filling tasks;
\item If possible, increase the number of evaluators, or reduce the number of questions per participant. In the above experiments each participant filled from 110 to 187 gaps, divided into small groups. Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}\comment{EA: the participants did not leave the gaps empty, they just did not attempt the next portions of evaluation after finishing some and getting bored. or did I misunderstand the use of captchas? }
\item To account for the adaptation period, provide training tasks before the main evaluations take place.
\end{enumerate}
